### 2013-09-18 intro.tex
page numbering, layouting...
trying to fix abstract :-/

\section{List of Contributions}
Contributions:
- Introduces general clustering algorithm that improves pipeline in serveral places.
- Shows that weakly supervised object class segmentation (not semantic segmentation) is possible.
- Provides general efficient framework for structured prediction.
- Analyze max-margin learning algorithms in different regimes.
- shows exact learning is possible in loopy graphs and improves results (sometimes?).
- shows importance of edge features.

>>>>

\section{List of Contributions}
Contributions:
\begin{itemize}
\item Introduces general clustering algorithm that improves pipeline in serveral places.
\item Shows that weakly supervised object class segmentation (not semantic segmentation) is possible.
\item Provides general efficient framework for structured prediction.
\item Analyze max-margin learning algorithms in different regimes.
\item shows exact learning is possible in loopy graphs and improves results (sometimes?).
\item shows importance of edge features.
\end{itemize}

### 2013-09-18 main.tex
fixes von hannes
page numbering, layouting...
trying to fix abstract :-/

\documentclass[12pt,toc=bibnumbered, a4paper,twoside,DIV=11,BCOR=.5cm]{scrbook}
%\documentclass[12pt,toc=bibnumbered, a4paper,twoside]{book}

@@

%\usepackage{garamondx}
%\usepackage[urw-garamond]{mathdesign}
\usepackage{xspace}
\onehalfspacing
%\KOMAoptions{DIV=last}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

@@

\begin{center}
\Large\sffamily
\vspace*{1cm}
\textbf{{\huge Learning Structured Prediction For Semantic Image Segmentation}}\\[5mm]
Dissertation\\%[3mm]
zur\\%[3mm]
Erlangung des Doktorgrades (Dr.\ rer.\ nat.)\\%[3mm]
der\\%[3mm]
Mathematisch-Naturwissenschaftlichen Fakult\"at\\%[3mm]
der\\%[3mm]
Rheinischen Friedrich-Wilhelm Universit\"at Bonn\\%[3mm]
vorgelegt von\\%[5mm]
{\LARGE Andreas Christian M\"uller}\\%[5mm]
aus\\%[3mm]
Offenbach am Main\\%[10mm]
Bonn, September 2013
\end{center}
\end{titlepage}
%
\thispagestyle{empty}
\vspace*{5cm}
\Large\sffamily
\begin{center}
Angefertigt mit Genehmigung\\
der Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der Rheinischen Friedrich-Wilhelms-Universit\"at Bonn\\
\end{center}
1. Gutachter Prof. Dr. Sven Behnke\\
2. Gutachter Prof. Dr. Reinhard Klein\\
Tag der Promotion:\\
Erscheinungsjahr: 2013\\
\end{spacing}
\pagenumbering{Roman}
\tableofcontents
%\listoffigures
%\listoftables
\chapter*{Zusammenfassung}
Automatische Segementierung und Erkennung von semantischen Klassen in
natuerlichen Bildern ist ein wichtiges offenes Problem des machinellen Sehens.
Einer der am weitest verbreiteten Ans\"atze zu dieser Aufgabe basiert auf
Conditional Random Fields und Verfahren der strukturierten Vorhersage. Die
Vorhersage f\"ur Modelle zur Bild-Markierung mittels Conditional Random Fields
ist f\"ur gew\"ohnlich nur approiximativ moeglich, bedingt d\"urch das
Vorhandensein von Kreisen in den zugrunde liegenden Nachbarschaftsgraphen.
Aus diesem Grund, und trotz der Beliebtheit von Conditional Random Field
Verfahren, ist das strukturierte Lernen f\"ur solche Anwendungen schlecht
verstanden.
%
Wir untersuchen verschiedene Lernalgorithmen in diesem
Zusammenhang. Wir zeigen, dass es m\"oglich ist trotz des Vorhandenseins von
Kreisen in den Nachbarschaftsgraphen exakte strukturierte Modelle zu lernen.
Wir stellen ausserdem eine Software-Bibliothek vor, die es erlaubt einen
weitreichenden Vergleich der besten Lernverfahren f\"ur strukturiertes Lernen
durchzuf\"uhren.
Unsere Studie erlaubt uns eine Charakterisation der betrachteten Algorithmen
in einer Reihe von Anwendungen, insbesondere der semantischen Segmentierung und
Objektklassensegmentierung.

\chapter*{Abstract}
Automatic segmentation and recognition of semantic classes in natural images is an important open problem in computer vision.
One of the most popular approaches to this task is based on conditional random fields and structured prediction.
For image labeling tasks, prediction using conditional random fields is usually only approximate, because of the presence
of loops of the neighborhood graph in the the image. Consequently, learning in
this setting is still little understood, despite the popularity of these
methods.
%
We investigate several learning algorithms in this setting. We show
how structured models can be learned exactly in practical settings, even in the
presence of many loops in the underlying neighborhood graphs.  We introduce a
software library that allows us to perform extensive empirical comparisons of
state-of-the-art structured learning approaches. This allows us to characterize
their practical properties in a range of applications, in particular for semantic segmentation
and object class segmentation.

@@

Last but not least I am greatful to Anna M\"uller for her support,
encouragement and company. Thank you for giving me the strength needed for this
endeavour.

\include{intro}

@@

\include{itm/paper}

\include{partial_supervision/paper}

@@

\chapter{Conclusion and Future Direction}
Current progress in computer vision \dots at an incredible speed,
with new methods and applications emerging constantly.
Still, there are large differences in the paradigms employed by different researchers.
While it is generally aknowledged that context is one of the most important cues
in recognizing objects and interpreting scence, there is little consensus on how
contextual information should be modelled, or even learned.
The structured prediction paradigm allows a principled approach to incoporating
predictions within an image, and to include context information.
We investigated the potential of structural methods for the application of semantic image segmentation
and object recognition, appications to which structured methods have been applied successfully before.
We focus on \emph{learning} of structural models and the interaction of inference and learning
in the neighborhood models typically employed for semantic segmentation.
The thesis presents a software implementation of a variety of popular learning algorithms for structural
support vector machines, together with a thorough evaluation of their properties in the presence
of approximate inference.
We demonstated that exact learning is possible, even in the presence of loops in the
\dots factor graph. We also demonstrate the power of conditioning interactions
on data, learning spacial interactions in a RGB-D setting.

There are several directions for future research that we think would be interesting to pursue as an extention
of the presented results: 
\paragraph[Inference Machines]
Recently \citet{stoyanov2011empirical} started a new trend in structured
prediction, which is sometimes called ``inference machines''. The basic
principle is simple: view the process of prediction with a given inference
procedure as a prediction method and optimize parameters of this method using
empirical risk minimization.  The work of \citet{stoyanov2011empirical}, used
loopy belief propagation as their inference algorithm and the optimization is
carried out simply using gradient descent on the non-convex but differentiable
loss function.  Other recent work in this direction includes
\citet{krahenbuhlparameter}, who used mean-field inference in a fully connected
conditional random field and \citet{jancsarylearning}, who used closed form
inference in a Gaussian CRF\@.
While these algorithms show great promise, their relation to the traditional
approach of structured prediction used in this work is mostly unclear. In particular,
if exact traditional learning is possible in a model, it is unclear how much
can be gained by direct empirical risk minimization.
Little empirical comparision was performed, and we are not aware of theoretical
work in this direction, leaving much room for future investigation of this
direction.
\paragraph[Non-Linear Models]
In this work, we only considered models that are linear in the input features--though
features are highly non-linear in the original input pixels.
Kernelization of structural support vector machines is trivial in theory, but
had only limited success in the context of CRFs for image
segmentation~\citep{lucchi2012structured}.
Two major alternatives for non-linear CRFs were proposed in the literature,
conditional neural fields~\citep{peng2009conditional} based on neural networks,
and decision tree fields~\citep{nowozin2011decision}, based on decision trees.
Conditional neural fields have only applied to sequence classification so far,
and extending them to our setting of semantic image segmentation would be very
interesting. Decision tree fields~(DTFs) on the other hand have been applied to loopy
graphs for image processing, but not for higher level tasks such as semantic
segmentation. If it is possible to include context in a meaningful way, it
might be possible to solve even object-centric tasks such as object class
segmentation with DTFs.

\paragraph[Latent Variable Models]
\paragraph[Feature Design]

%Vision is challenging
%context matters
%structured prediction allows modeling and learning of context and relations
%thesis gives insight into some relations of approximate inference and learning
%gives framework for evaluation
%comparison of methods

which low-level cues?
superpixels?
non-linear models?
latent variables?
inference engines

>>>>

\documentclass[12pt,toc=bibnumbered, a4paper,twoside,DIV=11,BCOR=.5cm]{scrbook}
%\documentclass[12pt,toc=bibnumbered, a4paper,twoside,DIV=calc]{scrbook}
%\documentclass[12pt,toc=bibnumbered, a4paper,twoside]{book}

@@

%\usepackage{garamondx}
%\usepackage[urw-garamond]{mathdesign}
\usepackage{xspace}
\onehalfspacing
\KOMAoptions{DIV=last}

\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}

@@

\begin{center}
\Large\sffamily
\vspace*{1cm}
\textbf{{\huge Learning Structured Prediction For Semantic Image Segmentation}}\\[5mm]
Dissertation\\%[3mm]
zur\\%[3mm]
Erlangung des Doktorgrades (Dr.\ rer.\ nat.)\\%[3mm]
der\\%[3mm]
Mathematisch-Naturwissenschaftlichen Fakult\"at\\%[3mm]
der\\%[3mm]
Rheinischen Friedrich-Wilhelm Universit\"at Bonn\\%[3mm]
vorgelegt von\\%[5mm]
{\LARGE Andreas Christian M\"uller}\\%[5mm]
aus\\%[3mm]
Offenbach am Main\\%[10mm]
Bonn, September 2013
\end{center}
\end{titlepage}
%
%\thispagestyle{empty}
\clearpage
\pagenumbering{Roman}
\setcounter{page}{2}
\vspace*{5cm}
\Large\sffamily
\begin{center}
Angefertigt mit Genehmigung\\
der Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der Rheinischen Friedrich-Wilhelms-Universit\"at Bonn\\
\end{center}
1. Gutachter Prof. Dr. Sven Behnke\\
2. Gutachter Prof. Dr. Reinhard Klein\\
Tag der Promotion:\\
Erscheinungsjahr: 2013\\
\end{spacing}
\tableofcontents
%\listoffigures
%\listoftables
\chapter*{Zusammenfassung}
Automatische Segementierung und Erkennung von semantischen Klassen in
nat\"urlichen Bildern ist ein wichtiges offenes Problem des machinellen Sehens.
Diese Arbeit untersucht drei m\"oglichen Ansi\"atzen der Erkennung:
ohne \"Uberwachung, mit \"Uberwachung auf Ebene von Bildern und mit \"Uberwachung auf Ebene
von Pixeln. Diese Arbeit setzt sich aus drei Teilen zusammen.

Im ersten Teil der Arbeit schlagen wir einen Clustering-Algorithmus vor,
der eine neuartige, informationstheoretische Zielfunktion optimiert. Wir zeigen, dass
der vorgestellte Algorithmus \"ublichen Standardverfahren aus der Literatur gegen\"uber
klare Vorteile auf vielen verschiedenen Datens\"atzen hat.

Der zweite Teil dieser Arbeit stellt ein Verfahren zur automatischen
Segmentierung und Erkennung von Objektklassen in nat\"urlichen Bildern vor, der
mit Hilfe von Supervision in Form von Klassen-Pr\"asenz auf Bildern in der Lage
ist ein Segmentierungsmodell zu Lernen. Nach unserem Wissen ist dies der erste
Ansatz dieser Art, der sich speziell auf komplexe Objektklassen bezieht.

Der dritte und letzte Teil er Arbeit untersucht einen der am weitest
verbreiteten Ans\"atze zu semantischen Segmentierung und
Objektklassensegmentierung, n\"amlich Conditional Random Fields verbunden mit Verfahren
der strukturierten Vorhersage.
%Die Vorhersage f\"ur Modelle zur Bild-Markierung
%mittels Conditional Random Fields ist f\"ur gew\"ohnlich nur approiximativ
%moeglich, bedingt d\"urch das Vorhandensein von Kreisen in den zugrunde
%liegenden Nachbarschaftsgraphen.  Aus diesem Grund, und trotz der Beliebtheit
%von Conditional Random Field Verfahren, ist das strukturierte Lernen f\"ur
%solche Anwendungen schlecht verstanden.
%
Wir untersuchen verschiedene Lernalgorithmen des struktuierten Lernens, insbesondere
im Zusammenhang mit approximativer Inferenz. Wir zeigen, dass es m\"oglich ist trotz des Vorhandenseins von
Kreisen in den Nachbarschaftsgraphen exakte strukturierte Modelle zur Bildsegmentierung zu lernen.
Wir stellen ausserdem eine Software-Bibliothek vor, die es erlaubt einen
weitreichenden Vergleich der besten Lernverfahren f\"ur strukturiertes Lernen
durchzuf\"uhren.
Unsere Studie erlaubt uns eine Charakterisation der betrachteten Algorithmen
in einer Reihe von Anwendungen, insbesondere der semantischen Segmentierung und
Objektklassensegmentierung.

\chapter*{Abstract}
Automatic segmentation and recognition of semantic classes in natural images is
an important open problem in computer vision.
This work investigates three different approaches to recognition: without supervision,
with supervision on level of images, and with supervision on the level of pixels.
The thesis comprises three parts.

The first part introduces a clustering algorithm, that optimizes a novel information-theoretic
objective function. We show that the proposed algorithm has clear advantages over standard algorithm
from the Literature on a wide array of datasets.

The second part of this work proposes an algorithm for automatic segmentation
and recognition of object classes in natural images, that learns a segmentation model solely from
annotation in form of presence and absence of object classes in images. To the best of our knowledge,
this is the first algorithm of this kind, that explicitly addresses complex object classes.

The third and last part of this work investigates one of the most popular
approaches to the task of object class segmentation and semantic segmentation,
based on conditional random fields and structured prediction.
%
%For image labeling tasks, prediction using conditional random fields is usually
%only approximate, because of the presence of loops of the neighborhood graph in
%the the image. Consequently, learning in this setting is still little
%understood, despite the popularity of these methods.
%
We investigate several learning algorithms, in particular in combination with
approximate inference procedures. We show how structured models for image
segmentation can be learned exactly in practical settings, even in the presence
of many loops in the underlying neighborhood graphs.  We introduce a software
library that allows us to perform extensive empirical comparisons of
state-of-the-art structured learning approaches. This allows us to characterize
their practical properties in a range of applications, in particular for
semantic segmentation and object class segmentation.

@@

Last but not least I am greateful to Anna M\"uller for her support,
encouragement and company. Thank you for giving me the strength needed for this
endeavour.
\cleardoublepage
\pagenumbering{arabic}
\include{intro}

@@

%\include{itm/paper}

%\include{partial_supervision/paper}

@@

\chapter{Conclusion and Future Direction}
Current research in computer vision progresses at an incredible speed, with new
methods and applications emerging constantly.  Still, there are large
differences in the paradigms employed by different researchers.  While it is
generally acknowledged that context is one of the most important cues in
recognizing objects and interpreting scenes, there is little consensus on how
contextual information should be modelled, or even learned.
The structured prediction paradigm allows a principled approach to
incorporating predictions within an image, and to include context information.
We investigated the potential of structural methods for the application of
semantic image segmentation and object recognition, applications to which
structured methods have been applied successfully before.  We focus on
\emph{learning} of structural models and the interaction of inference and
learning in the neighborhood models typically employed for semantic
segmentation.
The thesis presents a software implementation of a variety of popular learning
algorithms for structural support vector machines, together with a thorough
evaluation of their properties in the presence of approximate inference.
We demonstrate that exact learning is possible, even in the presence of loops in the
underlying factor graph. We also demonstrate the power of conditioning interactions
on data, learning spacial interactions in a RGB-D setting.

There are several directions for future research that we think would be
interesting to pursue as an extension of the presented results, in particular in
extending the second and third part:
\paragraph[Large-Scale Weakly Supervised Object Class Segmentation]
google
\paragraph[Inference Machines]
Recently \citet{stoyanov2011empirical} started a new trend in structured
prediction, which is sometimes called ``inference machines''. The basic
principle is simple: view the process of prediction with a given inference
procedure as a prediction method and optimize parameters of this method using
empirical risk minimization.  The work of \citet{stoyanov2011empirical}, used
loopy belief propagation as their inference algorithm and the optimization is
carried out simply using gradient descent on the non-convex but differentiable
loss function.  Other recent work in this direction includes
\citet{krahenbuhlparameter}, who used mean-field inference in a fully connected
conditional random field and \citet{jancsarylearning}, who used closed form
inference in a Gaussian CRF\@.
While these algorithms show great promise, their relation to the traditional
approach of structured prediction used in this work is mostly unclear. In particular,
if exact traditional learning is possible in a model, it is unclear how much
can be gained by direct empirical risk minimization.
Little empirical comparison was performed, and we are not aware of theoretical
work in this direction, leaving much room for future investigation of this
direction.
\paragraph[Non-Linear Models]
In this work, we only considered models that are linear in the input features--though
features are highly non-linear in the original input pixels.
Kernelization of structural support vector machines is trivial in theory, but
had only limited success in the context of CRFs for image
segmentation~\citep{lucchi2012structured}.
Two major alternatives for non-linear CRFs were proposed in the literature,
conditional neural fields~\citep{peng2009conditional} based on neural networks,
and decision tree fields~\citep{nowozin2011decision}, based on decision trees.
Conditional neural fields have only applied to sequence classification so far,
and extending them to our setting of semantic image segmentation would be very
interesting. Decision tree fields~(DTFs) on the other hand have been applied to loopy
graphs for image processing, but not for higher level tasks such as semantic
segmentation. If it is possible to include context in a meaningful way, it
might be possible to solve even object-centric tasks such as object class
segmentation with DTFs.

\paragraph[Latent Variable Models]
Something latent
\paragraph[Feature Design]
features are important
which low-level cues?
superpixels?
